---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
library(tidyverse)
theme_set(theme_bw())
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# xrf

[![Travis-CI Build Status](https://travis-ci.org/paleolimbot/xrf.svg?branch=master)](https://travis-ci.org/paleolimbot/xrf)

The goal of xrf is to provide tools to read, plot, and interpret X-Ray fluorescence spectra.

## Installation

You can install xrf from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("paleolimbot/xrf")
```

## Example

Read in a Panalytical XRF spectrum and plot it.

```{r example}
library(tidyverse)
library(xrf)

pan_example_dir <- system.file("spectra_files/Panalytical", package = "xrf")
pan_files <- list.files(pan_example_dir, ".mp2", full.names = TRUE)
specs <- read_xrf_panalytical(pan_files)
specs %>%
  unnest(.spectra) %>%
  ggplot(aes(x = energy_kev, y = cps, col = SampleIdent)) +
  geom_line() +
  facet_wrap(vars(ConditionSet), scales = "free_y")
```

## Baselines

The **xrf** package can use several existing methods for estimating "background" or "baseline" values. The most useful of these for XRF spectra is the Sensitive Nonlinear Iterative Peak (SNIP) method, implemented in the **Peaks** package.

```{r}
specs %>%
  slice(3) %>%
  xrf_add_baseline_snip(iterations = 20) %>%
  unnest() %>%
  filter(energy_kev <= 15) %>%
  ggplot(aes(x = energy_kev)) +
  geom_line(aes(y = cps, col = "raw")) +
  geom_line(aes(y = baseline, col = "baseline")) +
  geom_line(aes(y = cps - baseline, col = "cps - baseline"))
```

## Smoothing

```{r}
specs %>%
  slice(3) %>%
  xrf_add_baseline_snip(iterations = 20) %>%
  xrf_add_smooth_gaussian() %>%
  unnest() %>%
  filter(energy_kev <= 15) %>%
  ggplot(aes(x = energy_kev)) +
  geom_line(aes(y = cps, col = "raw"), alpha = 0.3) +
  geom_line(aes(y = smooth - baseline, col = "smooth"))
```

## Peaks

```{r, eval=FALSE}
energy_kev <- specs$.spectra[[1]]$energy_kev
spec <- specs$.spectra[[1]]$fit
peaks <- Peaks::SpectrumSearch(spec, background = TRUE, iterations = 15, threshold = .5)
ggplot(tibble(energy_kev, spec, deconv = peaks$y), aes(energy_kev)) +
  geom_line(aes(y = spec, col = "original")) +
  geom_line(aes(y = deconv, col = "deconv")) +
  geom_vline(xintercept = energy_kev[peaks$pos], alpha = 0.2, col = "red") +
  scale_y_sqrt()
```



```{r, eval=FALSE}
oreas22d <- specs %>%
  filter(SampleIdent == "oreas 22d") %>%
  unnest(.spectra) %>%
  mutate(cps = counts / LiveTime)

xrf_en <- xrf_energies %>%
  crossing(tibble(ConditionSet = unique(oreas22d$ConditionSet))) %>%
  group_by(ConditionSet) %>%
  mutate(
    data = list(oreas22d[oreas22d$ConditionSet == ConditionSet[1],]),
    counts = approx(data[[1]]$energy_kev, data[[1]]$counts, energy_kev)$y,
    fit = approx(data[[1]]$energy_kev, data[[1]]$fit, energy_kev)$y,
    background = approx(data[[1]]$energy_kev, data[[1]]$background, energy_kev)$y,
    cps = approx(data[[1]]$energy_kev, data[[1]]$cps, energy_kev)$y
  ) %>%
  select(-data)

library(plotly)
plot_ly() %>%
  add_lines(x = ~energy_kev, y = ~cps, color = ~ConditionSet, hoverinfo = "none", 
            data = oreas22d) %>%
  add_markers(x = ~energy_kev, y = ~cps, text = ~element, color = ~ConditionSet, data = xrf_en)
```


